Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-02 paper/main.tex
proof of Lemma 3 unfinished (klizi se mi oci, az rano)

\newcommand{\Nset}{{\mathbb{N}_0}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
%\def\Inf{\ensuremath{\mathsf{Inf}}}
%\def\Fin{\ensuremath{\mathsf{Fin}}}
\DeclareMathOperator{\Inf}{\mathsf{Inf}}
\DeclareMathOperator{\Fin}{\mathsf{Fin}}
\newcommand{\minf}{\mathit{inf}}
\newcommand{\mks}{\mathit{mks}}
%\newcommand{\oM}{\widehat{M}}
\newcommand{\oM}{N}
\newcommand{\rem}{\mathit{nm}}
\newcommand{\mcycle}{\mathit{cycle}}
\newcommand{\msat}{\mathit{satisfies}}
\newcommand{\Te}{T_{\vec{e}}}
\def\false{\mathit{false}}
\def\true{\mathit{true}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mB}{\mathcal{B}}

@@

A \emph{path} from a state $p$ to a state $q$ is a finite sequence of
adjacent transitions of the form
$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
is called a nontrivial \emph{strongly connected component (SCC)} if
for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
the rest of this paper, SCC always refers to a maximal SCC. Given an
SCC $S$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
denotes the set of all transitions between states in $S$. Further, for
each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
\mks(t)\}$. \todo{The sets $\delta_m$ are called \emph{acceptance
    sets}. Given an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we
  denote the set $\delta_{S,m}=\delta_S\cap\delta_m$.} A set of
transitions $T\subseteq\delta$ is called a \emph{cycle} is there
exists a path from a state $p$ to the same state containing all
transitions of $T$ and no transition outside $T$. Finally, we assume
that each TELA $\mA$ contains only states $q$ that are reachable from
the initial state $q_I$ (i.e., $q=q_I$ or there is a path from $q_I$
to $q$) as states that are not reachable from $q_I$ can be eliminated
without any impact on $L(\mA)$.

@@

\subsection{Strict version $\mcycle_3(\vec{e})$}
Before we give the definition of $\mcycle_3(\vec{e})$, we prove that
cycles can be characterised in the following way.
\begin{lemma}
  A set $T\subseteq\delta$ is a cycle if for each set of
  states $S\subseteq Q$ it holds that
  \[
    \begin{split}
    \textrm{if $T$ contains a transition leading from a state in $S$ to a state outside $S$,~~~~}\\[-.5ex]
    \textrm{then it also contains a transition leading from a state outside $S$ to a state in $S$.}
    \end{split}
    \tag{$3$}\label{eq:lem3} 
  \]  
\end{lemma}
\begin{proof}
  \todo{Ted to musim dokazat.}
\end{proof}

>>>>

\newcommand{\Nset}{{\mathbb{N}_0}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
%\def\Inf{\ensuremath{\mathsf{Inf}}}
%\def\Fin{\ensuremath{\mathsf{Fin}}}
\DeclareMathOperator{\Inf}{\mathsf{Inf}}
\DeclareMathOperator{\Fin}{\mathsf{Fin}}
\newcommand{\minf}{\mathit{inf}}
\newcommand{\mks}{\mathit{mks}}
%\newcommand{\oM}{\widehat{M}}
\newcommand{\oM}{N}
\newcommand{\rem}{\mathit{nm}}
\newcommand{\mcycle}{\mathit{cycle}}
\newcommand{\msat}{\mathit{satisfies}}
\newcommand{\Te}{T_{\vec{e}}}
\newcommand{\pre}{S_\mathit{pre}}
\newcommand{\post}{S_\mathit{post}}
\def\false{\mathit{false}}
\def\true{\mathit{true}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mB}{\mathcal{B}}

@@

A \emph{path} from a state $p$ to a state $q$ is a finite sequence of
adjacent transitions of the form
$(q_0,u_0,M_0,q_1)(q_1,u_1,M_1,q_2)\dots(q_{n-1},u_{n-1},M_{n-1},q_n)\in\delta^+$
such that $p=q_0$ and $q=q_n$. A nonempty set of states $S\subseteq Q$
is called a nontrivial \emph{strongly connected component (SCC)} if
for each $p,q\in S$ there is a path from $p$ to $q$. An SCC S is
\emph{maximal} if there is no SCC $S'$ satisfying $S\subsetneq S'$. In
the rest of this paper, SCC always refers to a maximal SCC. Given a
set of states $S\subseteq Q$, $\delta_S=\delta\cap (S\times \Sigma \times 2^M \times S)$
denotes the set of all transitions between states in $S$. Further, for
each mark $m\in M$, let $\delta_m$ denotes the set of all transitions
marked with $m$, i.e., $\delta_m=\{t\in\delta\mid m\in
\mks(t)\}$.
% \todo{The sets $\delta_m$ are called \emph{acceptance sets}. Given
% an SCC $S$ and a mark $m\in M$, by $\delta_{S,m}$ we denote the set
% $\delta_{S,m}=\delta_S\cap\delta_m$.}
A set of transitions $T\subseteq\delta$ is called a \emph{cycle} is
there exists a path from a state $p$ to the same state containing all
transitions of $T$ (potentially multiple times) and no transition
outside $T$. Finally, we assume that each TELA $\mA$ contains only
states $q$ that are reachable from the initial state $q_I$ (i.e.,
$q=q_I$ or there is a path from $q_I$ to $q$) as states that are not
reachable from $q_I$ can be eliminated without any impact on $L(\mA)$.

@@

\subsection{Strict version $\mcycle_3(\vec{e})$}
Before we give the definition of $\mcycle_3(\vec{e})$, we prove that
cycles can be characterised in the following way.
\begin{lemma}
  A nonempty set $T\subseteq\delta$ is a cycle if and only if for each set of
  states $S\subseteq Q$ it holds that
  \begin{enumerate}
  \item\label{en:1} all transitions in $T$ lead from a state in $S$ to a state in $S$ (i.e., $T\subseteq\delta_S$) or
  \item\label{en:2} all transitions in $T$ lead from a state outside $S$ to a state outside $S$ (i.e., $T\subseteq\delta_{Q\smallsetminus S}$) or
  \item\label{en:3} $T$ contains a transition leading from a state in $S$ to a state outside $S$ and a transition leading from a state outside
    $S$ to a state in $S$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  We first prove the direction ``$\Longrightarrow$''. Let $T$ be a
  cycle and $S\subseteq Q$ be an arbitrary set of states. We show that
  if (\ref{en:1}) and (\ref{en:2}) do not hold, then (\ref{en:3}) has
  to hold. Hence, assume that $T\not\subseteq\delta_S$ and
  $T\not\subseteq\delta_{Q\smallsetminus S}$. Then there are two
  cases.
  \begin{itemize}
  \item $T$ contains a transtition $t\in\delta_S$ and a transition
    $t'\in\delta_{Q\smallsetminus S}$. The definition of a cycle
    implies that there exists a path $t_1t_2\ldots t_n\in T^+$ from a
    state $p$ back to $p$ containing both $t$ and $t'$. However, this
    implies that $T$ contains a transtion leading from a state in $S$
    to a state outside $S$ and a transtion leading from a state outside $S$
    to a state in $S$.
  \item $T$ contains a transition $t$ leading from a state in $S$ to a
    state outside $S$ (or vice versa). However, as $T$ is a cycle,
    there exists a path $t_1t_2\ldots t_n\in T^+$ that leads from a
    state $p$ to the same state and contains $t$. Hence, $T$ has to
    contain also a transition leading from a state outside $S$ to a
    state in $S$ (or vice versa).
  \end{itemize}
  In both cases, $\ref{en:3}$ holds.

  Now we prove the opposite direction ``$\Longleftarrow$''. As $T$ is
  nonempty, there exists a transition $t\in T$ leading from a state $p$.
  We use this state to define the sets $\post,\pre\subseteq Q$ of states as 
  \begin{align}\notag
    \post&=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }p\textrm{ to }q\}\\
    \pre &=\{q\in Q\mid \textrm{there is a path in }T^+\textrm{ from }q\textrm{ to }p\}
  \end{align}
  Note that each transition $t'\in T$ leading from a state in
  $q\in\post$ leads also to a state in $\post$ as the path in $T^+$
  leading form $p$ to $q$ can be prolonged by this transition. In
  other words, there is no transtition in $T$ leading from a state
  in $\post$ to a state outside $\post$. Similarly, one can show that
  there is no transition in $T$ leadning from outside $\pre$ to $\pre$.

  \todo{Jan is here}

  us we
  choose Let us fix
  a transition $t\in T$ leading from a state $p$ to a state $q$. Let
  $\post$ be the set of all states reachable from $q$ using
  transitions in $T$ and $\pre$ be the set of all states from which

  that $T\subseteq\delta$ is a nonemtpy set such that~\eqref{eq:lem3}
  holds for each set $S\subset Q$.
\end{proof}
